---
title: "Assignment 4"
format:
  html:
    toc: true
editor: visual
---

***How to do it?***:

-   Open the Quarto file of this assignment ([link](assignment4q.qmd)) in Rstudio.

-   Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
library(dplyr)
```

------------------------------------------------------------------------

#### 1. Install `tidyverse` package

An R package can be installed by `install.packages` function. Install `tidyverse` if you have not done so.

```{r, eval=FALSE}
install.packages('dplyr')
install.packages('tidyverse')
```

------------------------------------------------------------------------

#### 2. Read the data using `read_csv`

Use `read_csv` function to import the US Covid 19 data at [link](https://bryantstats.github.io/math421/data/all-states-history.csv). Don't forget to import `tidyverse` (library(tidyverse)) so that you can use read_csv.

```{r}
library(tidyverse)
df = read_csv("C:/Users/student/OneDrive - Bryant University/Documents/GitHub/math421assigntments/all-states-history.csv")
```

------------------------------------------------------------------------

#### 3. Fix the date and create some new variables

`lubridate` is a package of the `tidyverse` packages. We will make uses of `lubridate` in this question.

-   Use the below codes to create `month`, `weekday` and `monthday` variables

```{r}
library(lubridate)
df$month = month(df$date)

# day of the week
df$weekday = wday(df$date)

# day of the month
df$monthday <- mday(df$date)
```

------------------------------------------------------------------------

#### 4. Create new variables with `case_when`.

The function `case_when` is a good option to create a new variable from existing variable. For example, this below codes create a new variable, `daily_death`, from `deathIncrease` variable. `deathIncrease` is the number of daily new death by Covid19. The new variable `daily_death` takes three values: low (if `deathIncrease` less than 3), medium (`deathIncrease` from 3 to 14), and high (`deathIncrease` more than 14). Please notice that this can also be done in a different way as shown in Assignment 3.

```{r, eval=FALSE}
df$daily_death <- case_when(
  df$deathIncrease <3 ~ 'low',
  df$deathIncrease <=14 ~ 'medium',
  TRUE ~ 'high'
)
```

-   Create variable `month2` that takes three values: early_month (day of the month from 1-10), mid_month (day of the month from 11-20), and end_month (day of the month \> 20).

    ```{r}
    df$month2 <- case_when(
      as.numeric(format(df$date, "%d")) <= 10 ~ "early_month",
      as.numeric(format(df$date, "%d")) <= 20 ~ "mid_month",
      TRUE ~ "end_month"
    )
    ```

-   Create variable `weekend` that takes two values: 1 if it's Saturday or Sunday and 0 otherwise.

    ```{r}
    df$weekend <- case_when(
      weekdays(df$date) %in% c("Saturday", "Sunday") ~ 1,
      TRUE ~ 0
    )
    ```

------------------------------------------------------------------------

#### 5. Select function

Use the select function to deselect the column `totalTestsViral` from the data.

```{r}
df <- select(df, -totalTestsViral)
```

------------------------------------------------------------------------

#### 6. Pipe Operator ( %\>% )

Pipe operator offers another way to write R codes. Many times, it makes the codes more readable. Pipe works very well with all the `tidyverse` packages. Refer to [these slides](https://bryantstats.github.io/math421/slides/4_dplyr.html#15) (slide 15, 16, 17 and 18) to rewrite the below codes using pipe operator

```{r}
x <- c(1:10)

# square root of x
x %>% sqrt

# sum of the square root of x
x %>% sqrt %>% sum

#log of the sums of the square roots of x
x %>% sqrt %>% sum %>% log

# log base 2 of 16
16 %>% log(2)

```

------------------------------------------------------------------------

#### 7. Combo 1: group_by + summarise

This combo is used when you want to apply a function/calculation to different groups of the data. For example, to calculate the average number of cases (`positiveIncrease`) by `dataQualityGrade`, we use:

`{r, eval=FALSE} #df %>%  #  group_by(weekday) %>%  # summarise(mean(positiveIncrease))`

```         
```

-   Calculate the median number of cases (`positiveIncrease`) by `month`

```{r}
df %>%
  group_by(df$month) %>%
  summarise(median_cases = median(positiveIncrease))
```

-   Calculate the average number of cases (`positiveIncrease`) by `month2`

```{r}
df %>%
  group_by(month2) %>%
  summarise(mean_cases = mean(positiveIncrease))
```

-   Calculate the median number of cases (`positiveIncrease`) by `weekend`

    ```{r}
    df %>%
      group_by(df$weekend) %>%
      summarise(median_cases = median(positiveIncrease))
    ```

------------------------------------------------------------------------

#### 8. Combo 2: filter + group_by + summarise

An example: to calculate the average number of cases (`positiveIncrease`) in January and February separately, we use:

```{r, eval=FALSE}
df %>% 
  filter(month == 1|month == 2) %>% 
  group_by(month) %>% 
  summarise(positve_increase = mean(positiveIncrease))
```

-   Calculate the median number of cases (`positiveIncrease`) on the weekend by `month` in October and November 2020.

    ```{r}
    df %>%
      filter(
        weekend == 1,
        month %in% c(10, 11),
        format(date, "%Y") == "2020"
      ) %>%
      group_by(month) %>%
      summarise(
        median_cases = median(positiveIncrease)
      )
    ```

-   Calculate the average number of death at different periods of a month (`month2` variable) in Fall 2020

    ```{r}
    df %>%
      filter(
        month %in% c(9, 10, 11),
        format(date, "%Y") == "2020"
      ) %>%
      group_by(month2) %>%
      summarise(
        mean_deaths = mean(death, is.na = TRUE)
      )
    ```

-   Compare the average number of hospitalizations between weekdays and weekends in Summer 2020

    ```{r}
    df %>%
      filter(
        format(date, "%Y") == "2020",
        month %in% c(6, 7, 8)
      ) %>%
      group_by(weekend) %>%
      summarise(
        mean_hospitalized = mean(hospitalizedIncrease, is.na = TRUE)
      )
    ```

-   Redo Questions 14 and 15 in [Assignment 3](assignment3q.html) using the combos. Notice: you also need to use the data used in Assignment 3.

------------------------------------------------------------------------

#### 9. Combo 3: filter + group_by + summarise + arrange

Use the arrange function to find a month that has the highest number of deaths on the weekend.

```{r}
df %>%
  filter(weekend == 1) %>%
  group_by(month) %>%
  summarise(mean_deaths = mean(death, is.na = TRUE)) %>%
  arrange(desc(mean_deaths))
```

------------------------------------------------------------------------

#### 10. Use your own dataset and implement the follows functions or combos.

You can use the [Adult Census Income](https://bryantstats.github.io/math421/data/adult_census.csv) or [Titanic data](https://bryantstats.github.io/math421/data/titanic.csv).

```{r}
titanic = read_csv("C:/Users/student/OneDrive - Bryant University/Documents/GitHub/math421assigntments/titanic.csv")
```

using the titanic data

-   select

    ```{r}
    titanic %>% select(Name, Sex, Age, Survived)
    ```

-   filter

    ```{r}
    titanic %>% filter(Survived == 1)
    ```

-   mutate

    ```{r}
    titanic %>% mutate(Age_group = ifelse(Age < 18, "Child", "Adult"))
    ```

-   summarise

    ```{r}
    titanic %>% summarise(Mean_age = mean(Age, na.rm = TRUE))
    ```

-   arrange

    ```{r}
    titanic %>% arrange(Sex)
    ```

-   count

    ```{r}
    titanic %>% count(Sex)
    ```

-   count + arrange

    ```{r}
    titanic %>% count(Pclass) %>% arrange(desc(n))
    ```

-   filter + count + arrange

    ```{r}
    titanic %>% filter(Survived == 1) %>% count(Pclass) %>% arrange(desc(n))
    ```

-   group_by + summarise

    ```{r}
    titanic %>% group_by(Pclass) %>% summarise(Mean_age = mean(Age, na.rm = TRUE))
    ```

-   filter + group_by + summarise

    ```{r}
    titanic %>% filter(Survived == 1) %>% group_by(Pclass) %>% summarise(Mean_age = mean(Age, na.rm = TRUE))
    ```

-   filter + group_by + summarise + arrange

    ```{r}
    titanic %>% filter(Survived == 1) %>% group_by(Pclass) %>% summarise(Mean_age = mean(Age, na.rm = TRUE))
    ```
