---
title: "Assignment 4"
format:
  html:
    toc: true
editor: visual
---

***How to do it?***:

-   Open the Quarto file of this assignment ([link](assignment4q.qmd)) in Rstudio.

-   Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas.


```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
library(dplyr)
```

------------------------------------------------------------------------

#### 1. Install `tidyverse` package

An R package can be installed by `install.packages` function. Install `tidyverse` if you have not done so.

```{r, eval=FALSE}
install.packages('dplyr')
install.packages('tidyverse')
```

------------------------------------------------------------------------

#### 2. Read the data using `read_csv`

Use `read_csv` function to import the US Covid 19 data at [link](https://bryantstats.github.io/math421/data/all-states-history.csv). Don't forget to import `tidyverse` (library(tidyverse)) so that you can use read_csv.

------------------------------------------------------------------------

#### 3. Fix the date and create some new variables

`lubridate` is a package of the `tidyverse` packages. We will make uses of `lubridate` in this question.

-   Use the below codes to create `month`, `weekday` and `monthday` variables

```{r, eval=FALSE}
library(lubridate)
df$month = month(df$date)

# day of the week
df$weekday = wday(df$date)

# day of the month
df$monthday <- mday(df$date)
```

------------------------------------------------------------------------

#### 4. Create new variables with `case_when`.

The function `case_when` is a good option to create a new variable from existing variable. For example, this below codes create a new variable, `daily_death`, from `deathIncrease` variable. `deathIncrease` is the number of daily new death by Covid19. The new variable `daily_death` takes three values: low (if `deathIncrease` less than 3), medium (`deathIncrease` from 3 to 14), and high (`deathIncrease` more than 14). Please notice that this can also be done in a different way as shown in Assignment 3.

```{r, eval=FALSE}
df$daily_death <- case_when(
  df$deathIncrease <3 ~ 'low',
  df$deathIncrease <=14 ~ 'medium',
  TRUE ~ 'high'
)
```

-   Create variable `month2` that takes three values: early_month (day of the month from 1-10), mid_month (day of the month from 11-20), and end_month (day of the month \> 20).

-   Create variable `weekend` that takes two values: 1 if it's Saturday or Sunday and 0 otherwise.

------------------------------------------------------------------------

#### 5. Select function

Use the select function to deselect the column `totalTestsViral` from the data.

df \<- select(df, -totalTestsViral)

------------------------------------------------------------------------

#### 6. Pipe Operator ( %\>% )

Pipe operator offers another way to write R codes. Many times, it makes the codes more readable. Pipe works very well with all the `tidyverse` packages. Refer to [these slides](https://bryantstats.github.io/math421/slides/4_dplyr.html#15) (slide 15, 16, 17 and 18) to rewrite the below codes using pipe operator

```{r}
x <- c(1:10)

# square root of x
x %>% sqrt

x %>% sqrt %>% sum

x %>% sqrt %>% sum %>% log

# log base 2 of 16
16 %>% log(2)

```

------------------------------------------------------------------------

#### 7. Combo 1: group_by + summarise

This combo is used when you want to apply a function/calculation to different groups of the data. For example, to calculate the average number of cases (`positiveIncrease`) by `dataQualityGrade`, we use:

###```{r, eval=FALSE}
#df %>% 
#  group_by(weekday) %>% 
# summarise(mean(positiveIncrease))
```

-   Calculate the median number of cases (`positiveIncrease`) by `month`

#df %\>% group_by(month) %\>% summarise(median_cases = median(positiveIncrease))

-   Calculate the average number of cases (`positiveIncrease`) by `month2`

   df %\>% group_by(month2) %\>% summarise(mean_cases = mean(positiveIncrease))

-   Calculate the median number of cases (`positiveIncrease`) by `weekend`

    df %\>% group_by(weekend) %\>% summarise(median_cases = median(positiveIncrease))

------------------------------------------------------------------------

#### 8. Combo 2: filter + group_by + summarise

An example: to calculate the average number of cases (`positiveIncrease`) in January and February separately, we use:

```{r, eval=FALSE}
df %>% 
  filter(month==1|month==2) %>% 
  group_by(month) %>% 
  summarise(positve_increase = mean(positiveIncrease))
```

-   Calculate the median number of cases (`positiveIncrease`) on the weekend by `month` in October and November 2020.

    df %\>% filter(weekend == "weekend" & month %in% c(10, 11) & year == 2020) %\>% group_by(month) %\>% summarise(median_cases = median(positiveIncrease))

-   Calculate the average number of death at different periods of a month (`month2` variable) in Fall 2020

    df %\>% filter(month %in% c(9, 10, 11) & year == 2020) %\>% group_by(month2) %\>% summarise(mean_deaths = mean(death))

-   Compare the average number of hospitalizations between weekdays and weekends in Summer 2020

    df %\>% filter(month %in% c(6, 7, 8) & year == 2020) %\>% group_by(weekend) %\>% summarise(mean_hospitalizations = mean(hospitalizedIncrease))

-   Redo Questions 14 and 15 in [Assignment 3](assignment3q.html) using the combos. Notice: you also need to use the data used in Assignment 3.

    14: df %\>% filter(STATE == "Rhode Island" & format(DATE, "%Y") == "2021") %\>% summarise(mean_cases = mean(CASES))

    15: df %\>% filter(STATE == "Rhode Island" & format(DATE, "%Y") == "2021") %\>% group_by(weekdays) %\>% summarise(median_cases = median(CASES))

------------------------------------------------------------------------

#### 9. Combo 3: filter + group_by + summarise + arrange

Use the arrange function to find a month that has the highest number of deaths on the weekend.

df %\>% filter(weekday %in% c("Saturday", "Sunday")) %\>% group_by(month) %\>% summarise(mean_deaths = mean(death)) %\>% arrange(desc(mean_deaths))

------------------------------------------------------------------------

#### 10. Use your own dataset and implement the follows functions or combos.

You can use the [Adult Census Income](https://bryantstats.github.io/math421/data/adult_census.csv) or [Titanic data](https://bryantstats.github.io/math421/data/titanic.csv).

using the titanic data

-   select

    titanic %\>% select(Name, Sex, Age, Survived)

-   filter

    titanic %\>% filter(Survived == 1)

-   mutate

    titanic %\>% mutate(Age_group = ifelse(Age \< 18, "Child", "Adult"))

-   summarise

    titanic %\>% summarise(Mean_age = mean(Age, na.rm = TRUE))

-   arrange

    titanic %\>% arrange(Age)

-   count

    titanic %\>% count(Sex)

-   count + arrange

    titanic %\>% count(Pclass) %\>% arrange(desc(n))

-   filter + count + arrange

    titanic %\>% filter(Survived == 1) %\>% count(Pclass) %\>% arrange(desc(n))

-   group_by + summarise

    titanic %\>% group_by(Pclass) %\>% summarise(Mean_age = mean(Age, na.rm = TRUE))

-   filter + group_by + summarise

    titanic %\>% filter(Survived == 1) %\>% group_by(Pclass) %\>% summarise(Mean_age = mean(Age, na.rm = TRUE))

-   filter + group_by + summarise + arrange

    titanic %\>% filter(Survived == 1) %\>% group_by(Pclass) %\>% summarise(Mean_age = mean(Age, na.rm = TRUE))
