---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Part 1"
---

***How to do it?***:

-   Open the Rmarkdown file of this assignment ([link](assignment5_part1.Rmd)) in Rstudio.

-   Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

------------------------------------------------------------------------

**The data:**

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

------------------------------------------------------------------------

**Questions**

1.  Use `read_csv` to import the data. Show the number of NAs for each columns.

    ```{r}
    library(tidyverse)
    adult <- read_csv("adult_census_missing.csv")
    sum(is.na(adult))
    ```

2.  Using function `aggr` the VIM package to plot the number of NAs for each column.

    ```{r}
    #library(VIM)

    #aggr(adult,
         #numbers = TRUE,
         #prop = TRUE)
    ```

3.  Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data.

    ```{r}
    table(adult$workclass)
    table(adult$occupation)
    table(adult$native.country)
    ```

4.  Replace all the forms of missing values found to NA

    ```{r}
    adult[adult == "?"] <- NA
    adult[adult == "Unknown"] <- NA

    colSums(is.na(adult))
    ```

5.  Replot the number of NAs for each column.

    ```{r}
    #aggr(adult)
    ```

6.  Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs). How many rows left after removing?

    ```{r}
    adult_no_na <- na.omit(adult)

    nrow(adult_no_na)
    ```

7.  Approach 2 to handle NAs: Fill all the NAs with the previous or next value. (Hint: Using `fill` function). Save the dataset after filling as a different data. The original data is unchanged (still have NAs).

    ```{r}
    library(dplyr)

    adult_filled <- adult %>%
      fill(everything(), .direction = "downup")

    colSums(is.na(adult_filled))
    ```

8.  Approach 3 to handle NAs: For numeric variable, replace the NAs by the median. For categorical variables, replace the NAs to the majority.

    ```{r}
    adult_median <- adult

    adult_median$age[is.na(adult_median$age)] <- median(adult_median$age, na.rm = TRUE)

    adult_median$fnlwgt[is.na(adult_median$fnlwgt)] <- median(adult_median$fnlwgt, na.rm = TRUE)

    adult_median$education.num[is.na(adult_median$education.num)] <- median(adult_median$education.num, na.rm = TRUE)

    adult_median$capital.gain[is.na(adult_median$capital.gain)] <- median(adult_median$capital.gain, na.rm = TRUE)

    adult_median$capital.loss[is.na(adult_median$capital.loss)] <- median(adult_median$capital.loss, na.rm = TRUE)

    adult_median$hours.per.week[is.na(adult_median$hours.per.week)] <- median(adult_median$hours.per.week, na.rm = TRUE)

    get_mode <- function(x) {
      x <- x[!is.na(x)]
      names(sort(table(x), decreasing = TRUE))[1]
    }
    adult_median$workclass[is.na(adult_median$workclass)] <- get_mode(adult_median$workclass)

    adult_median$education[is.na(adult_median$education)] <- get_mode(adult_median$education)

    adult_median$marital.status[is.na(adult_median$marital.status)] <- get_mode(adult_median$marital.status)

    adult_median$occupation[is.na(adult_median$occupation)] <- get_mode(adult_median$occupation)

    adult_median$relationship[is.na(adult_median$relationship)] <- get_mode(adult_median$relationship)

    adult_median$race[is.na(adult_median$race)] <- get_mode(adult_median$race)

    adult_median$sex[is.na(adult_median$sex)] <- get_mode(adult_median$sex)

    adult_median$native.country[is.na(adult_median$native.country)] <- get_mode(adult_median$native.country)

    adult_median$income[is.na(adult_median$income)] <- get_mode(adult_median$income)

    colSums(is.na(adult_median))
    ```
